\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}

\title{HeteroKV: Source-Aware Approximate Long-Context KV Cache for 128K Generation under Fixed HBM Budgets}
\author{Anonymous}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Long-context generation with full KV caching scales memory linearly with
sequence length, making 128K-token contexts difficult to support on
24GB-class GPUs for 7B-scale instruction models. We present HeteroKV, an
approximate long-context cache that keeps a bounded active KV set in GPU HBM
and stores evicted KV in compressed DRAM-side records. HeteroKV combines sink
tokens, tail tokens, heavy-hitter retention, compressed DRAM storage, and
source-aware token-level retrieval. It is not a lossless reconstruction of
native full attention; instead, it targets stable memory use, functional
generation, and controlled semantic loss under a fixed HBM budget.

On Qwen2.5-7B-Instruct at 128K context length, the promoted source-aware NIAH
configuration achieves 24/24 accuracy across depths 25\%, 50\%, 75\%, and
90\% on an A100 server constrained to a 4090-like memory envelope. The run uses
a 22 GiB PyTorch cap and a 30 GiB own-process fuse. Mean decode latency is
98.12 ms/step, or 1.88x a wide-memory A100 FullKV reference. A FullKV 128K
negative control fails under the same 22 GiB cap with CUDA OOM. SourceCopy-
disabled PPL runs show +1.20\%, +1.66\%, +0.45\%, and +3.14\% relative deltas
on WikiText-2 suffix setups and +1.09\% on an IMDb 16K suffix setup. These
results support HeteroKV as a practical approximate cache for long-context
survival, while true RTX 4090 latency remains future hardware validation.
\end{abstract}

\section{Introduction}

Full KV caching is simple and accurate, but it scales linearly with context
length. At 128K tokens, a 7B model can exceed the usable memory budget of a
24GB GPU. HeteroKV addresses this gap by keeping a bounded active KV set in HBM
and moving evicted tokens into compressed DRAM-side storage. The design is an
Approximate Long-Context Cache rather than a lossless full-KV replacement.

\section{Method}

HeteroKV keeps sink tokens, tail tokens, heavy-hitter tokens, and selected
retrieved tokens in HBM. Evicted tokens are stored outside HBM in compressed
records. During decode, source-aware token-level retrieval can bring relevant
chunks back into the active cache. The strongest NIAH path uses source-aware
prefiltering and late-layer retrieval. Pure Query-Key dot-product retrieval is
reported separately as a negative control.

\section{Experimental Setup}

The primary model is Qwen2.5-7B-Instruct. The main context length is 128K. The
hardware claim is A100 under a 4090-like memory envelope: 22 GiB PyTorch cap
and 30 GiB own-process fuse. This does not claim native RTX 4090 latency.

\section{Results}

\subsection{128K NIAH}

The promoted source-aware path reaches 24/24 across required depths 25\%,
50\%, 75\%, and 90\%, using three seeds and two trials per depth. Mean decode
latency is 98.12 ms/step; the wide-memory A100 FullKV reference is
52.25 ms/step.

\subsection{Memory}

The parsed seed6004 memory log contains 64 prefill chunks. Active HBM KV reaches
8192 tokens and remains flat. DRAM compressed KV grows to 122880 tokens. Max
torch reserved memory is 21.33 GiB, and max nvidia-smi process memory is
21.82 GiB.

A FullKV 128K negative control under the same 22 GiB PyTorch cap fails with
CUDA OOM when PyTorch reports 20.90 GiB process memory in use against a
22.00 GiB allowance. The monitor does not trigger the 30 GiB safety fuse.

\subsection{PPL}

SourceCopy-disabled PPL remains separate from the source-aware NIAH path. The
validated suffix setups are:

\begin{center}
\begin{tabular}{lrrr}
\toprule
Dataset / window & FullKV PPL & HeteroKV PPL & Delta \\
\midrule
WikiText-2 14K & 2.9706 & 3.0063 & +1.20\% \\
WikiText-2 16K & 4.9896 & 5.0723 & +1.66\% \\
WikiText-2 16K offset32768 & 6.2955 & 6.3237 & +0.45\% \\
WikiText-2 32K & 6.5289 & 6.7336 & +3.14\% \\
IMDb 16K & 15.2031 & 15.3683 & +1.09\% \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Negative Controls}

A clean current pure dot-product top8/qhist64 128K control scores 0/4 on the
25\%/50\% subset, with mean decode 1005.04 ms/step and peak process memory
21.8242 GiB. Earlier pure-dot variants score 0/4 to 1/4. A broader pure
dot-product scaling diagnostic, run without source-aware filtering or
SourceCopy under the same 22 GiB cap, scores 4/8 at 16K, 5/8 at 32K, and
2/8 at 64K across required depths 25\%, 50\%, 75\%, and 90\%. This confirms
that pure QK scoring stayed within the memory envelope but did not provide the
quality of the promoted source-aware path.

\section{Limitations}

The strongest result is source-aware and should not be described as pure KV-only
dot-product retrieval. A100-under-cap measurements do not prove native RTX 4090
latency. PPL evidence is suffix-style evaluation through 32K plus IMDb 16K, not
true 128K PPL. The current 0\% NIAH template is non-discriminative because
FullKV also fails it.

\section{Conclusion}

HeteroKV demonstrates that an approximate long-context cache can support 128K
generation for a 7B model under a 24GB-class memory envelope while preserving
task accuracy in source-aware NIAH and keeping PPL degradation controlled in
SourceCopy-disabled evaluations.

\end{document}
